Review of Assigned Readings

### 2018-01-05 paper/paper.tex
final review of paper

%************************** Abstract ************************** 
\begin{abstract}
\pagenumbering{Roman}
Based on previous research in the stock market, this paper uses a Twitter sentiment analysis to determine the dynamics between the amount of Twitter activity regarding Bitcoin and the market price of Bitcoin. The analysis is based on two different data sets. Twitter activity has been measured by the number of tweets containing the term "Bitcoin" while Bitcoin prices were measured in the Bitcoin price index. The Vector Autoregression (VAR) model was used for the statistical analysis and the interpretation of the results.\\

@@

\subsection{Methodology}
\label{sec:Methodology}
To answer the above mentioned research question, a scientific approach has been applied. The paper consists out of three parts. First, the relevance of the topic is pointed out by the introduction, the research question, methodology and the scope (see chapter \ref{sec:intro}). Secondly, the related theory has been analysed to provide the reader with the necessary background information by briefly introducing both Bitcoin and Twitter in context of price prediction and behavioral finance. The covered research is based on forward research in regard to the stock market and cryptocurrencies with a special focus on sentiment analysis and Bitcoin \ref{sec:Background}. Further on,  Vector Autoregression (VAR) approach is introduced and its constrains in context of the paper are being outlined (see chapter \ref{sec:EconometricModellingandResults}). The third part contains the discussion of the results. The computation is based on a data set, which was gathered by Python scripts for the time period December 20th - 27th. Since the results are based on the observation of a short period, they have to be interpreted carefully. The data set consists of two data sources with the following characteristics: (1) Twitter tweets which contain the buzzword "Bitcoin" and (2) those containing historical Bitcoin prices. The two sources have been aggregated by using Python. The output of the aggregation was further processed with STATA. The third part of the paper is discussed in the chapter \ref{sec:EconometricModellingandResults} followed by a brief conclusion in chapter \ref{sec:Conclustion}.

@@

In relation to Bitcoin, researchers already showed that analysing investors sentiments can lead to reasonable correlations between Twitter sentiments and prices of cryptocurrencies \textcite[p.~7]{stenqvist2017predicting} , by either using a \textit{polarity classification} {-} classifying the language in tweets as either positive or negative (see \cite{colianni2015algorithmic}) {-} or a \textit{lexicon based approach} {-} attributes predefined words to specific sentiment classifications (see \cite{stenqvist2017predicting}). Hence, this paper choses a different focus area by limiting its scope to the number of daily tweets that contain the term “Bitcoin” without analysing investors interests any further. Based on analysis of the influence of Twitter bullishness on the stock market by \textcite{mao2015quantifying}, this paper further elaborates on the concept by applying the approach to Bitcoin prices.\\

\subsection{Data Collection}
For our data, we collected two separate time series datasets. One showing the Twitter activity regarding Bitcoin, with the second one showing the historical Bitcoin price development. Both datasets were collected hourly for the period from the 19.12.2017 starting at 20:00:00 UTC and ending on the 28.12.2017 at 10:00:00 UTC. We chose the hourly basis to conduct statistical analysis with a limited amount of observations as demanded by the feasibility of such a complex project. For the Twitter data we aggregated by each hour all tweets mentioning "Bitcoin". We used prices at the last second of every hour to represent the hourly Bitcoin price.\\
The following figure displays the Twitter activity with regard to Bitcoin for the observed time period. The first plot shows the hourly change of the aggregate number of tweets containing "Bitcoin". In the second and third plots we respectively did a first difference and a logarithmic first difference transformation of our main variable, in case it is needed in the later analysis. The graphs show overall a cyclical movement around 1200 tweets per hour and a single outburst befalling on 22.12.2017 between 15:00:01 and 21:00:00, with highest value being 4622 tweets in one hour.\\

@@

%************************** Chapter 3 - Econometric Modelling and Results **************************
\section{Econometric Modelling and Results}
\label{sec:EconometricModellingandResults}
As stated in the beginning, our purpose is to examine the underlying relationship between the number of tweets containing "Bitcoin" and Bitcoin price fluctuations. From our data collection, we are left with two time series datasets representing two variables. A broadly used statistical method to simultaneously analyze multiple time series is the Vector Autoregression (VAR) approach. In this method, the endogenous variables are determined both by their own historical values and by the historical values of the other endogenous variables (\cite[pp.~4--5]{luetkepohl2007new}).\\
To generate the following econometric results we used the statistical software STATA.

@@

%************************** Chapter 4 - Conclusion **************************
\section{Conclusion}
\label{sec:Conclustion}
Understanding, describing and predicting the current phenomenon of cryptocurrencies has been the objective of many researchers. A variety of methods and approaches have been applied und mixed results have been produced. Our goal for this seminar paper was to investigate the possibility of the existence of a correlation between Twitter activity regarding Bitcoin and the price evolution of Bitcoin. After collecting hourly data for approximately 8 days we generated 2 time-series datasets and decided to perform a statistical analysis by using the Vector Autoregression (VAR) approach.\\

According to our regression results, changes in Twitter activity due to more or less tweets about Bitcoin have a very small and insignificant effect on future changes of Bitcoin price. The validity of this statement was also verified and confirmed by the Granger causality test. At this point, however, it must be pointed out that only a small period of time has been considered in this analysis and the results cannot be generalized for the overall trend of cryptocurrencies in the recent time or for the future. We are also aware that we may have observed a period of abnormal behavior of Bitcoin prices compared to the trend of the last months.\\

We encourage further research in the same direction as this paper and would be interested to see an analysis with observations over a longer period.

>>>>

%************************** Abstract ************************** 
\begin{abstract}
\pagenumbering{Roman}
Based on previous research in the stock market, this paper uses a Twitter sentiment analysis to determine the dynamics between the amount of Twitter activity regarding Bitcoin and the market price of Bitcoin. The analysis is based on two different data sets. Twitter activity has been measured by the number of tweets containing the term "bitcoin" while Bitcoin prices were measured in the Bitcoin price index. The Vector Autoregression (VAR) model was used for the statistical analysis and the interpretation of the results.\\

@@

\subsection{Methodology}
\label{sec:Methodology}
To answer the above mentioned research question, a scientific approach has been applied. The paper consists out of three parts. First, the relevance of the topic is pointed out by the introduction, the research question, methodology and the scope (see chapter \ref{sec:intro}). Secondly, the related theory has been analysed to provide the reader with the necessary background information by briefly introducing both Bitcoin and Twitter in context of price prediction and behavioral finance. The covered research is based on forward research in regard to the stock market and cryptocurrencies with a special focus on sentiment analysis and Bitcoin \ref{sec:Background}. Further on,  Vector Autoregression (VAR) approach is introduced and its constrains in context of the paper are being outlined (see chapter \ref{sec:EconometricModellingandResults}). The third part contains the discussion of the results. The computation is based on a data set, which was gathered by Python scripts for the time period December 20th - 27th. Since the results are based on the observation of a short period, they have to be interpreted carefully. The data set consists of two data sources with the following characteristics: (1) Twitter tweets which contain the buzzword "bitcoin" and (2) those containing historical Bitcoin prices. The two sources have been aggregated by using Python. The output of the aggregation was further processed with STATA. The third part of the paper is discussed in the chapter \ref{sec:EconometricModellingandResults} followed by a brief conclusion in chapter \ref{sec:Conclustion}.

@@

In relation to Bitcoin, researchers already showed that analysing investors sentiments can lead to reasonable correlations between Twitter sentiments and prices of cryptocurrencies \textcite[p.~7]{stenqvist2017predicting} , by either using a \textit{polarity classification} {-} classifying the language in tweets as either positive or negative (see \cite{colianni2015algorithmic}) {-} or a \textit{lexicon based approach} {-} attributes predefined words to specific sentiment classifications (see \cite{stenqvist2017predicting}). Hence, this paper choses a different focus area by limiting its scope to the number of daily tweets that contain the term “bitcoin” without analysing investors interests any further. Based on analysis of the influence of Twitter bullishness on the stock market by \textcite{mao2015quantifying}, this paper further elaborates on the concept by applying the approach to Bitcoin prices.\\

\subsection{Data Collection}
For our data, we collected two separate time series datasets. One showing the Twitter activity regarding Bitcoin, with the second one showing the historical Bitcoin price development. Both datasets were collected hourly for the period from the 19.12.2017 starting at 20:00:00 UTC and ending on the 28.12.2017 at 10:00:00 UTC. We chose the hourly basis to conduct statistical analysis with a limited amount of observations as demanded by the feasibility of such a complex project. For the Twitter data we aggregated by each hour all tweets mentioning "bitcoin". We used prices at the last second of every hour to represent the hourly Bitcoin price.\\
The following figure displays the Twitter activity with regard to Bitcoin for the observed time period. The first plot shows the hourly change of the aggregate number of tweets containing "bitcoin". In the second and third plots we respectively did a first difference and a logarithmic first difference transformation of our main variable, in case it is needed in the later analysis. The graphs show overall a cyclical movement around 1200 tweets per hour and a single outburst befalling on 22.12.2017 between 15:00:01 and 21:00:00, with highest value being 4622 tweets in one hour.\\

@@

%************************** Chapter 3 - Econometric Modelling and Results **************************
\section{Econometric Modelling and Results}
\label{sec:EconometricModellingandResults}
As stated in the beginning, our purpose is to examine the underlying relationship between the number of tweets containing "bitcoin" and Bitcoin price fluctuations. From our data collection, we are left with two time series datasets representing two variables. A broadly used statistical method to simultaneously analyze multiple time series is the Vector Autoregression (VAR) approach. In this method, the endogenous variables are determined both by their own historical values and by the historical values of the other endogenous variables (\cite[pp.~4--5]{luetkepohl2007new}).\\
To generate the following econometric results we used the statistical software STATA.

@@

%************************** Chapter 4 - Conclusion **************************
\section{Conclusion}
\label{sec:Conclustion}
Understanding, describing and predicting the current phenomenon of cryptocurrencies has been the objective of many researchers. A variety of methods and approaches have been applied and mixed results have been produced. Our goal for this project was to investigate the possibility of the existence of a correlation between Twitter activity regarding Bitcoin and the price evolution of Bitcoin. After collecting hourly data for 8 days, we generated two time-series datasets and decided to perform a statistical analysis by using the Vector Autoregression (VAR) approach.\\

According to our regression results, changes in Twitter activity caused by a higher or lower number of tweets about Bitcoin, have a very small and insignificant effect on future changes of Bitcoin price. The validity of this statement was also verified and confirmed by the Granger causality test. At this point, however, it must be pointed out that only a small period of time has been considered in this analysis and the results cannot be generalized for the overall trend of cryptocurrencies in the recent time or for the future. We are also aware, that we may have observed a period of abnormal behavior of Bitcoin prices compared to the trend of the last months.\\

We encourage further research in the same direction as this paper could serve as a blueprint together with the accompanying process documentation and would be interested to see an analysis with observations over a longer period.
